Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-01-11 Draft/abstract.tex
added EPJC template

\begin{abstract}
\noindent
Abstract to be written here.
\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

\tableofcontents
\newpage

>>>>

\begin{abstract}
\noindent

In this article, a detailed study of the vector-boson scattering for two positively charged W bosons is presented.
In particular, a comparison between the full next-to-leading (NLO) QCD corrections against several approximations is carried out.
This study is not only performed in the usual fiducial region used by experimental collaborations but also in a more inclusive set-up.
This allows to infer precisely the quality of such approximations.
Finally, NLO predictions matched to various parton shower are discussed.
This study allows thus to infer the systematic errors related to vector-boson scattering at the NLO-QCD level and beyond.

\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

%\tableofcontents
\newpage

### 2018-01-11 Draft/definition.tex
added EPJC template

Definition of the process. \\
Explaining the various contributions. \\
Show few Feynman diagrams and explain that it possesses also tri-boson contributions in the s-channel. \\
This is explained in some details in Ref.~\cite{Biedermann:2017bss}.

\begin{figure}[t]
\begin{center}
          \includegraphics[width=0.30\linewidth]{feynman/LO_EW_5}
          \raisebox{.5ex}{\includegraphics[width=0.35\linewidth]{feynman/LO_EW_2}}
          \raisebox{-1.8ex}{\includegraphics[width=0.32\linewidth]{feynman/LO_EW_3}}
\end{center}
        \caption{Sample tree-level diagrams that contribute to the process $\Pp\Pp\to\mu^+\nu_\mu\Pe^+\nu_{\Pe}\Pj\Pj$ at order $\mathcal{O}{\left(\alpha^{6}\right)}$.
        In addition to typical VBS contribution (left), this order also possesses $s$-channel (middle) and tri-boson contributions (right). }
\label{diag:LO}
\end{figure}

>>>>

The scattering of two positively charge W bosons is proceeding at the LHC through the partonic process:
%
\begin{equation}
 {\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}.
\end{equation}

This process possesses three LO contributions of different orders.
The first one is of order $\mathcal{O}{\left(\alpha^{6}\right)}$ is referred to as EW contributions.
In addition to typical VBS contributions as shown on the left of Fig.~\ref{diag:LO}, it also features $s$-channel contributions.
These type of contributions will play a particular role in the study of the various contributions in Section \ref{subsec:contributions}.
Some of them take the form of decay chains as for example the diagram represented in the middle of Fig.~\ref{diag:LO} while others are tri-boson contributions (right of Fig.~\ref{diag:LO})
The VBS diagrams typically dominate the full process.
But all these contributions form a single gauge-invariant set of contributions and therefore cannot be separated.

The process can also be mediated via a gluon connecting the two quark lines while the W bosons are radiated from the quark lines.
These contributions are of order $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$ and usually feature different kinematic behaviours than the EW contribution.
Nonetheless they share the same final state and therefore constitute an irreducible background.

Finally, due to the specific colour structure of these two classes of amplitudes, there exists non-zero interferences.
These are of order $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$ and are usually small but not negligible for realistic experimental set-ups \cite{Biedermann:2017bss}.

Usually, in experimental measurements, special VBS-cuts are designed in order to enhance the EW contribution over the QCD one.
These cuts are based on the fact that the two contributions have rather different kinematic.
The EW contribution is characterised by two jets with large rapidity in the peripheral region as well as a large invariant mass.
The two W bosons are mostly produced centrally.
This is in opposition with the QCD component which favour jets in the central region.
Therefore, the event selection usually involve rapidity-difference and invariance mass cuts for the jets.
This will also be discussed in Section \ref{subsec:contributions}.

\begin{figure*}[t]
\begin{center}
          \includegraphics[width=0.30\linewidth]{feynman/LO_EW_5}
          \raisebox{.5ex}{\includegraphics[width=0.35\linewidth]{feynman/LO_EW_2}}
          \raisebox{-1.8ex}{\includegraphics[width=0.32\linewidth]{feynman/LO_EW_3}}
\end{center}
        \caption{Sample tree-level diagrams that contribute to the process $\Pp\Pp\to\mu^+\nu_\mu\Pe^+\nu_{\Pe}\Pj\Pj$ at order $\mathcal{O}{\left(\alpha^{6}\right)}$.
        In addition to typical VBS contribution (left), this order also possesses $s$-channel (middle) and tri-boson contributions (right). }
\label{diag:LO}
\end{figure*}

### 2018-01-11 Draft/description.tex

  \item {{\sc{Phantom}}~\cite{Ballestrero:2007xq} is a dedicated tree--level Monte Carlo for six parton final states at $pp,\, p\bar{p}$
  and $e^+e^-$ colliders at $\alpha_{ew}^6$ and $\alpha_{ew}^4\alphas^2$ including interferences between the two sets of diagrams.
It employs complete tree--level matrix elements in the complex--mass scheme~\cite{Denner:2006ic} computed via the modular helicity 
formalism~\cite{Ballestrero:1999md,Ballestrero:1994jn}. The integration uses a multichannel approach~\cite{BERENDS1985441} and an adaptive
strategy~\cite{PETERLEPAGE1978192}. {\sc{Phantom}} generates unweighted events at parton--level for both the SM and a few instances of BSM theories.}

>>>>

  \item {{\sc{Phantom}}~\cite{Ballestrero:2007xq} is a dedicated tree--level Monte Carlo for six parton final states at $pp,\, p\bar{p}$
  and $e^+e^-$ colliders at $\alpha^6$ and $\alpha^4\alphas^2$ including interferences between the two sets of diagrams.
It employs complete tree--level matrix elements in the complex--mass scheme~\cite{Denner:2006ic} computed via the modular helicity 
formalism~\cite{Ballestrero:1999md,Ballestrero:1994jn}. The integration uses a multichannel approach~\cite{BERENDS1985441} and an adaptive
strategy~\cite{PETERLEPAGE1978192}. {\sc{Phantom}} generates unweighted events at parton--level for both the SM and a few instances of BSM theories.}

### 2018-01-11 Draft/introduction.tex

The vector-boson scattering (VBS) process with two positively charged W boson is just starting to be measured at the LHC~\cite{Aad:2014zda,Aaboud:2016ffv,Khachatryan:2014sta}.
A whole class of new processes will therefore be measured during Run~II at the Large Hadron Collider (LHC).
For now the measurements of VBS processes are limited by statistics but the situation will change in a near future.
On the theoretical side, it is thus of prime importance to provide precise predictions and infer their related systematics.

The $\PW^+\PW^+$ scattering is probably the simplest VBS processes to compute and to measure due to its particular charge structure and low background, respectively.
Therefore, it is an ideal example for a detailed study of various theoretical predictions.
In the last few years, several (N)LO computations become available for both the VBS process~\cite{Jager:2009xx,Jager:2011ms,Denner:2012dz,Rauch:2016pai} and its QCD-induced irreducible background
process~\cite{Melia:2010bm,Melia:2011gk,Campanario:2013gea,Baglio:2014uba,Rauch:2016pai}.
These computations all relied on approximations while recently the complete NLO corrections became available \cite{Biedermann:2017bss}.
It is therefore interesting to infer in details the quality of the various approximations.
Indeed, apart from Ref.~\cite{Biedermann:2017bss} \MP{more references?} no detailed comparison of the VBS approximations have been carried out.

The hadronic process is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ and it posses three contributions at leading order (LO) 
[$\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$, and $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$].
They are refer to as EW, interference, and QCD contributions respectively.
The EW contribution is sometimes referred to as the VBS contribution even it possesses not-VBS contributions.
Therefore, we start with a LO study of these contributions as a function of typical VBS cuts.
This allows to understand the various contributions to the final state $\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$.
This is followed by a LO comparison between the various predictions at the level of the cross section and differential distributions.

At NLO, the process possesses four contributions of orders 
$\mathcal{O}{\left(\alpha^{7}\right)}$,
$\mathcal{O}{\left(\alphas\alpha^{6}\right)}$,
$\mathcal{O}{\left(\alphas^{2}\alpha^{5}\right)}$, and
$\mathcal{O}{\left(\alphas^{3}\alpha^{4}\right)}$.
The largest one is the EW corrections of order $\mathcal{O}{\left(\alpha^{7}\right)}$ \cite{Biedermann:2016yds,Biedermann:2017bss}.
The contribution of order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ is the second largest NLO contribution.
It is often referred as the QCD corrections to the VBS process.
In the following, this order is the one where our comparisons are focused on.
It is therefore usually referred to as simply \emph{NLO}.
As for the LO study, the various predictions are compared at the level of the cross section and differential distributions.

Finally, several predictions featuring parton shower are compared.
This allows to infer systematics differences between the various predictions.
This is the first time in the literature that such an analysis has been carried out \MP{True?}.

@@

The article is organised as follow:
in the first part (section~\ref{sec:definition}), the process studied is defined.
Then, various approximation at LO and NLO are described in section~\ref{sec:details}.
It is followed by a presentation of the various programs used for the simulations.
Sections~\ref{sec:LO} and \ref{sec:NLO} are devoted to a LO and NLO study, respectively.
Section~\ref{sec:matching} deals with the matching to parton shower.
The last section (section \ref{sec:conclusion}) consists in concluding remarks and recommendations.

>>>>

The vector-boson scattering (VBS) process with two positively charged W boson is now starting to be measured at the LHC~\cite{Aad:2014zda,Aaboud:2016ffv,Khachatryan:2014sta}.
A whole class of new processes will therefore be measured during Run~II at the Large Hadron Collider (LHC).
For now the measurements of VBS processes are limited by statistics but the situation will change in a near future.
On the theoretical side, it is thus of prime importance to provide precise predictions and infer their related systematic errors.

The $\PW^+\PW^+$ scattering is the simplest VBS processes to compute theoretically and to measure experimentally due to its particular charge structure and low background, respectively.
Therefore, it is an ideal example for a detailed study of various theoretical predictions.
In the last few years, several leading order (LO) and next-to-leading order (NLO) computations become available for both the VBS process~\cite{Jager:2009xx,Jager:2011ms,Denner:2012dz,Rauch:2016pai} and its QCD-induced irreducible background
process~\cite{Melia:2010bm,Melia:2011gk,Campanario:2013gea,Baglio:2014uba,Rauch:2016pai}.
These computations all relied on approximations while recently the complete NLO corrections became available \cite{Biedermann:2017bss}.
It is therefore interesting to infer in details the quality of the various approximations.
Indeed, apart from Ref.~\cite{Biedermann:2017bss} where it is commented on, \MP{more references?} no detailed comparison of the VBS approximations have been carried out.

The hadronic process is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ and it posses three contributions at leading order.
The orders of the contributions are $\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$, and $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$.
They are refer to as electroweak (EW), interference, and QCD contributions respectively.
\footnote{The EW contribution is sometimes referred to as the VBS contribution even it possesses non-VBS contributions.}
Therefore, the present work starts with a LO study of these three contributions as a function of typical VBS cuts.
This allows to identify the various contributions to the final state $\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$.
This is followed by a LO comparison between the various predictions at the level of the cross section and differential distributions.

At NLO, the process possesses four contributions of orders 
$\mathcal{O}{\left(\alpha^{7}\right)}$,
$\mathcal{O}{\left(\alphas\alpha^{6}\right)}$,
$\mathcal{O}{\left(\alphas^{2}\alpha^{5}\right)}$, and
$\mathcal{O}{\left(\alphas^{3}\alpha^{4}\right)}$.
The largest one is the EW corrections of order $\mathcal{O}{\left(\alpha^{7}\right)}$ \cite{Biedermann:2016yds,Biedermann:2017bss}.
The contribution of order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ is the second largest NLO contribution.
It is often referred to as the QCD corrections to the VBS process.
In the following, this order is the one where our comparisons are focused on.
It is therefore usually referred to as simply \emph{NLO}.
As for the LO study, the various predictions are compared at the level of the cross section and differential distributions.
In particular, this allows to infer the accuracy of the so-called VBS approximation.
To our knowledge, such a detailed study was still missing.

Finally, several predictions featuring parton shower are compared.
This allows to infer systematic differences between the various predictions.
This is the first time in the literature that such an analysis has been carried out \MP{True?}.

@@

The article is organised as follow:
in the first part (section~\ref{sec:definition}), the process studied is defined.
Then, various approximation at LO and NLO are described in section~\ref{sec:details}.
It is followed by a presentation of the various programs used for the simulations.
Sections~\ref{sec:LO} and \ref{sec:NLO} are devoted to a LO and NLO study, respectively.
Section~\ref{sec:matching} deals with the matching to parton shower.
The last section (section \ref{sec:conclusion}) consists in concluding remarks and recommendations for experimental collaborations.

### 2018-01-11 Draft/scanNLO.tex

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.39]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}
FIGURE
\caption{Cross section (fb) per bin of $(M_{jj},\,\Delta y_{jj})$ at NLO QCD $\mathcal{O}(\alpha^6\alphas)$, without any cut on the $jj$ pair kinematics. Results of \texttt{XXX} calculations.}\label{fig:mjjdyjj_2d_NLO}
\end{figure}

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_tu.pdf}
%\includegraphics[scale=0.395]{figures/scanfigures/.pdf}
\caption{Cross sections (fb) per bin of $(M_{jj},\,\Delta y_{jj})$ at $\mathcal{O}(\alpha^6\alphas)$, without any cut on the $jj$ pair kinematics: ratio of approximated squared amplitudes over the full matrix element. The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$ (left) and $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$ (right). Results of \texttt{VBFNLO} (approximated) and \texttt{RECOLA} (full) calculations. RIGHT SIDE FIG. IN PREPARATION. QUESTION: is RECOLA calc. 'exact' ?}\label{fig:ratio2d_NLO}
\end{figure}

>>>>

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.39]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}
FIGURE
\caption{Cross section (fb) per bin of $(M_{jj},\,\Delta y_{jj})$ at NLO QCD $\mathcal{O}(\alpha^6\alphas)$, without any cut on the $jj$ pair kinematics:  ratio of approximated squared amplitudes over the full matrix element. The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$. Results of \texttt{VBFNLO} (approximated) and \texttt{RECOLA} (full) calculations.}\label{fig:mjjdyjj_2d_NLO}
\end{figure}

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_tu.pdf}
%\includegraphics[scale=0.395]{figures/scanfigures/.pdf}
\caption{Cross sections (fb) per bin of $(M_{jj},\,\Delta y_{jj})$ at NLO QCD $\mathcal{O}(\alpha^6\alphas)$, without any cut on the $jj$ pair kinematics: ratio of approximated squared amplitudes over the full matrix element. The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$. Results of \texttt{VBFNLO} (approximated) and \texttt{RECOLA} (full) calculations.}\label{fig:ratio2d_NLO}
\end{figure}
